Global adjoint tomography -- Model GLAD-M25

### 2018-10-19 main.tex
Q&A

To determine the source parameters in the starting model,
we used the CMT inversion algorithm of \cite{liu2004spectral}.
This algorithm combines a normalized waveform difference misfit with an envelope difference misfit.
Specifically, the algorithm minimizes the misfit function
\begin{equation}
   \begin{split}
      \Phi =  \sum\limits_{c=1}^{C} \omega_c \sum\limits_{r=1}^{R_c} \omega_{cr}
       \sum\limits_{w=1}^{W_{cr}} \omega_{crw}\,
          & \left\{ \lambda\, \frac
              { \int \big[ d_w(t) - s_w(t - \Delta t) \big]^2 \mathrm{d}t}
              {\int \big[ d_w(t) \big]^2  \mathrm{d}t} \right.
       \\ & \quad \left. \mbox{} + (1 - \lambda)\, \frac
              {\int \big[ e(d_w(t)) - e(s_w(t - \Delta t)) \big]^2 \mathrm{d}t}
              {\int \big[ e(d_w(t)) \big]^2\mathrm{d}t} \right\}
              \quad ,
   \end{split}
\end{equation}
where~$d_w(t)$ denotes data in time window~$w$\,,
and~$s_w(t - \Delta t)$ the corresponding synthetic for model~GLAD-M15.
We allow the synthetics to shift relative to the data by an amount~$\Delta t$
--- effectively a station correction --- determined by cross correlation.
The envelope function is denoted by~$e(\,\cdot\,)$\,,
and the parameter~$\lambda$ determines the balance between fitting waveforms versus fitting envelopes.
The number of measurement categories is denoted by~$C$\,.
Following~\cite{ekstrom2012global},
seismograms were filtered between 50~s and 100~s
to select three-component body-wave windows,
and between 60~s and 100~s to select three-component
surface-wave windows,
resulting in six measurement categories, $c=1,\ldots,C$\,.
To balance their contributions,
each category is weighted by the reciprocal of the number of measurements in that
category, $\omega_c$\,.
The number of windows for a given receiver~$r$ in a given measurement category~$c$ is~$W_{cr}$\,.
Each such window is weighted equally, i.e.,~$\omega_{crw}=1$\,.
The number of receivers that records data in category~$c$ is denoted by~$R_c$\,.
As shown in Fig.~\ref{fig:source_correction}b,
the receivers are unevenly distributed across the globe,
with several dense arrays in the Northern Hemisphere and poor coverage in the Southern Hemisphere.
To balance the coverage,
we assign receiver weights~$\omega_{cr}$ based on the expression
\begin{equation}
\omega_{cr}^{-1} = N_c\,\sum_{r'=1}^{R_c} \exp\left[\mbox{}-\left(\frac{\Delta_{rr'}}{\Delta_c}\right)^2\right]
\quad ,
\label{eq:spatial_weights}
\end{equation}
{\color{Red} Wenjie: is it $\omega_{cr}^{-2}$ or $\omega_{cr}^{-1}$? And does the normalization factor~$N_c$ depend on~$c$?\\ }
where~$N_c$ is a normalization factor,
and where~$\Delta_{rr'}$ denotes the angular distance between receivers~$r$ and~$r'$\,.
The reference angular distance~$\Delta_c$ needs to be chosen such that the condition 
number of the diagonal weighting matrix defined by Eqn.~(\ref{eq:spatial_weights}) is not too large.
The calculation of the weights may be abstracted as: given a spatial distribution of
points on the unit sphere, determine a weighting associated with each point.

@@

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/misfit.pdf}
  \caption{Evolution of phase misfits from GLAD-M15 to GLAD-M25.
  Each color denotes the different stages of inversions, between
  which we make changes to the inversion by either tuning the
  weighting strategies or adding more data.
  (a) Evolution of the total phase misfit. (b) - (j) Misfit in
  each period band and component. (e) - (g) From GLAD-M18,
  we future split the 40-100 sec period band into two
  measurement categories: body(red diamond) waves and surface waves
  (red square).
  {\color{red} Wenjie, why do the graphs overlap at iteration 22, but not at iteration 18? Shouldn't 22 look like 18, i.e., no overlapping points?}
  }
  \label{fig:misfit}
\end{figure}

@@

{\color{red} Wenjie: Make sure the variance reductions are calculated as ln (Phi\_M25 / Phi\_M15).}

\begin{table}[!htb]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ~          &  Vertical (\%) & Radial (\%) &  Transverse (\%) \\
    \hline
    17--40~s  body waves    &   17.4 (31.1)   &       21.1 (37.1) &       24.6 (42.2) \\
    40--100~s body waves    &   16.8 (31.3)  &       21.3 (40.2)  &       21.7 (42.0) \\
    40--100~s surface waves &   28.5 (50.1)  &       28.3 (51.1) &       28.4 (53.8)  \\
    90--250~s surface waves &   14.5 (36.2)  &       13.4 (38.8) &       25.7 (38.6) \\
    \hline
  \end{tabular}\\
  \caption{Changes in fit in the twelve measurement categories
between the new model, GLAD-M25, and its starting model, GLAD-M15, and, in parentheses, between GLAD-M25 and model S362ANI combined with CRUST2.0, which was the starting model for the GLAD-M15 inversion.}
  \label{table:misfit_reduction_M15_M25}
\end{table}

>>>>

To determine the source parameters in the starting model,
we used the CMT inversion algorithm of \cite{liu2004spectral}.
This algorithm combines a normalized waveform difference misfit with an envelope difference misfit.
Specifically, the algorithm minimizes the misfit function
\begin{equation}
   \begin{split}
      \Phi =  \sum\limits_{c=1}^{C} \omega_c \sum\limits_{r=1}^{R_c} \omega_{cr}
       \sum\limits_{w=1}^{W_{cr}} \omega_{crw}\,
          & \left\{ \lambda\, \frac
              { \int \big[ d_w(t) - s_w(t - \Delta t) \big]^2 \mathrm{d}t}
              {\int \big[ d_w(t) \big]^2  \mathrm{d}t} \right.
       \\ & \quad \left. \mbox{} + (1 - \lambda)\, \frac
              {\int \big[ e(d_w(t)) - e(s_w(t - \Delta t)) \big]^2 \mathrm{d}t}
              {\int \big[ e(d_w(t)) \big]^2\mathrm{d}t} \right\}
              \quad ,
   \end{split}
\end{equation}
where~$d_w(t)$ denotes data in time window~$w$\,,
and~$s_w(t - \Delta t)$ the corresponding synthetic for model~GLAD-M15.
We allow the synthetics to shift relative to the data by an amount~$\Delta t$
--- effectively a station correction --- determined by cross correlation.
The envelope function is denoted by~$e(\,\cdot\,)$\,,
and the parameter~$\lambda$ determines the balance between fitting waveforms versus fitting envelopes.
The number of measurement categories is denoted by~$C$\,.
Following~\cite{ekstrom2012global},
seismograms were filtered between 50~s and 100~s
to select three-component body-wave windows,
and between 60~s and 100~s to select three-component
surface-wave windows,
resulting in six measurement categories, $c=1,\ldots,C$\,.
To balance their contributions,
each category is weighted by the reciprocal of the number of measurements in that
category, $\omega_c$\,.
The number of windows for a given receiver~$r$ in a given measurement category~$c$ is~$W_{cr}$\,.
Each such window is weighted equally, i.e.,~$\omega_{crw}=1$\,.
The number of receivers that records data in category~$c$ is denoted by~$R_c$\,.
As shown in Fig.~\ref{fig:source_correction}b,
the receivers are unevenly distributed across the globe,
with several dense arrays in the Northern Hemisphere and poor coverage in the Southern Hemisphere.
To balance the coverage,
we assign receiver weights~$\omega_{cr}$ based on the expression
\begin{equation}
\omega_{cr}^{-1} = N_c\,\sum_{r'=1}^{R_c} \exp\left[\mbox{}-\left(\frac{\Delta_{rr'}}{\Delta_c}\right)^2\right]
\quad ,
\label{eq:spatial_weights}
\end{equation}

{\color{Red} Wenjie: is it $\omega_{cr}^{-2}$ or $\omega_{cr}^{-1}$? And does the normalization factor~$N_c$ depend on~$c$?\\ }
{\color{Red} To Jeroen: I checked the code. we are using $\omega_{cr}^{-1}$. The normalization factor depends on c.}

where~$N_c$ is a normalization factor,
and where~$\Delta_{rr'}$ denotes the angular distance between receivers~$r$ and~$r'$\,.
The reference angular distance~$\Delta_c$ needs to be chosen such that the condition 
number of the diagonal weighting matrix defined by Eqn.~(\ref{eq:spatial_weights}) is not too large.
The calculation of the weights may be abstracted as: given a spatial distribution of
points on the unit sphere, determine a weighting associated with each point.

@@

{\color{red} To Jeroen: You are right. $\Delta$ is different for a given source and category. Actually, we don't directly control $\Delta$. Instead, we control the condition number of weights. So the $\Delta$ will vary with receiver distributions. }

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/misfit.pdf}
  \caption{Evolution of phase misfits from GLAD-M15 to GLAD-M25.
  Each color denotes the different stages of inversions, between
  which we make changes to the inversion by either tuning the
  weighting strategies or adding more data.
  (a) Evolution of the total phase misfit. (b) - (j) Misfit in
  each period band and component. (e) - (g) From GLAD-M18,
  we future split the 40-100 sec period band into two
  measurement categories: body(red diamond) waves and surface waves
  (red square).
  {\color{red} Wenjie, why do the graphs overlap at iteration 22, but not at iteration 18? Shouldn't 22 look like 18, i.e., no overlapping points?    \textbf{To Jeroen}: overlap means I calculate the misfit twice using the same mode. For example, in iteration M22, I caculated the misfit using the 1040 earthquakes and 1480 earthquakes, respectively. In M18, I haven't done so. If you want, I can calculate the misfit values and adding the overlap misfit point.}
  }
  \label{fig:misfit}
\end{figure}

@@

{\color{red} Wenjie: Make sure the variance reductions are calculated as ln (Phi\_M25 / Phi\_M15).}
{\color{red} Wenjie: I was using (Phi\_M25 - Phi\_M15) / Phi\_M15. So let me modify it?}
\begin{table}[!htb]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ~          &  Vertical (\%) & Radial (\%) &  Transverse (\%) \\
    \hline
    17--40~s  body waves    &   17.4 (31.1)   &       21.1 (37.1) &       24.6 (42.2) \\
    40--100~s body waves    &   16.8 (31.3)  &       21.3 (40.2)  &       21.7 (42.0) \\
    40--100~s surface waves &   28.5 (50.1)  &       28.3 (51.1) &       28.4 (53.8)  \\
    90--250~s surface waves &   14.5 (36.2)  &       13.4 (38.8) &       25.7 (38.6) \\
    \hline
  \end{tabular}\\
  \caption{Changes in fit in the twelve measurement categories
between the new model, GLAD-M25, and its starting model, GLAD-M15, and, in parentheses, between GLAD-M25 and model S362ANI combined with CRUST2.0, which was the starting model for the GLAD-M15 inversion.}
  \label{table:misfit_reduction_M15_M25}
\end{table}
